---
title: "Web_Downloading_Chilean_Climate_Data"
output: html_document
date: "2023-03-20"
---
This function will download weather data from metochile.gob. The data downloads minute data, and you can select by month. This is still a work in progress, future updates might include selecting by day. 

NOTE: r libraries lubridate and tidyverse are required for this.

Common weather variables:

Rainfall is AguaCaida
Cloud cover is Nubosidad
Pressure and Humidity is PresionHumedad
Temperature is Temperatura
Wind is Viento
Radiation is RadiacionGlobal

Copy and Paste exactly as they are in Spanish into 'weather_variable'

The different variables and weather stations can be found here:
https://climatologia.meteochile.gob.cl/application/index/menuTematicoEmas

Select the date by month in the format: yyyy/mm/dd

Here is an example:
data <- download_weather_data("2019/02/01", "2019/04/30", "390006", "AguaCaida")


```{r}
wind <- download_weather_data("2020/02/13", "2020/04/16", "390006", "Viento")
temp <- download_weather_data("2020/02/13", "2020/04/16", "390006", "Temperatura")
pres_hum <- download_weather_data("2020/02/13", "2020/04/16", "390006", "PresionHumedad")
```


```{r}
library(lubridate)
library(tidyverse)
library(plyr)

download_weather_data <- function(start_date, end_date, StationID, weather_variable, outDir = getwd()) {

  # Here we turn the start date and end date into the format that is used in the url used to download
  Months <- (interval(ymd(start_date), ymd(end_date)) %/% months(1))+1
  range <- seq(ymd(start_date), by = "month", length.out = Months) #finds the # of months of data to download
  range_zip <- mapply(c, year(range), month(range), SIMPLIFY = FALSE) #creates a list of years and months
  vec <- unlist(range_zip, recursive = TRUE, use.names = FALSE) #unlists the list
  
  # This will combine the year and month to a single number of 6 characters.
  YearMonth <- character(length(vec)/2)
  for (i in seq_along(YearMonth)) {
    YearMonth[i] <- paste0(vec[(2*i-1)], sprintf("%02d", vec[2*i])) #This adds a leading 0 if the month is less than 10 as well as creates a value of yyyymm which is used in the url
  }

  # Set a base url that will be modified with the modified start date, modified end date, station id, and weather variable
  url_base <- "https://climatologia.meteochile.gob.cl/application/datos/getDatosEma/StationID/StationID_Date_WeatherVariable.csv.zip"

  # Create an empty list to store the downloaded file names
  downloaded_files <- list()
  
  # Loop through all the dates in YearMonth and paste the station id and weather variable in the appropriate spots in the URL 
  for (date in YearMonth) {
    zipF <- paste(StationID,"_",date,".zip", sep = "")
    url_1 <- gsub("Date", date, url_base) #these lines sub the variables into the url
    url_2 <- gsub("WeatherVariable", weather_variable, url_1)
    url <- gsub("StationID", StationID, url_2) #this is the final url used to download the data
    download.file(url, zipF)
    file <- unzip(zipF, exdir=outDir)
    downloaded_files <- c(downloaded_files, file) # Add the downloaded files to the list
    file.remove(zipF) #removes the zip file to keep directory clean and tidy
  }

  # Create a vector of file names
  file_list <- grep(paste(YearMonth, collapse = "|"), downloaded_files, value = TRUE)

  # Read in the files into a list
  data_list <- lapply(file_list, function(x) {
    read.csv(x, sep = ";")
  })
  
  #use that function
  data_list <- drop_extra_cols(data_list)

  # Combine the data frames into one
 all_data <- rbind.fill(data_list)
  
  # Delete the downloaded .csv to clean up the directory
  data_list <- lapply(file_list, function(x) {
    file.remove(x)
  })
  
  #select only the data from the original date ranges.
  all_data <- filter(all_data, momento >= ymd(start_date) & momento <= ymd(end_date))

  # Return the combined data frame
  return(all_data)
}

``` 


